---
title: "Introduction to automated text analysis"
author: Pablo Barbera, Alex Hanna
date: March 30, 2016
output: html_document
---

In this session we will cover automated text analysis. We will start with basic string manipulation with R.

Our running example will be a random sample of 10,000 tweets mentioning hashtags related to international courts. We'll save the text of these tweets as a vector called `text'

```{r}
tweets <- read.csv("../datasets/icourts-tweets.csv", stringsAsFactors=F)
head(tweets)
text <- tweets$text
```

R stores the basic string in a character vector. `length` gets the number of items in the vector, while `nchar` is the number of characters in the vector.

```{r}
length(text)
text[1]
nchar(text[1])
```

Note that we can work with multiple strings at once.

```{r}
nchar(text[1:10])
sum(nchar(text[1:10]))
```

As we will see later, it is often convenient to convert all words to lowercase or uppercase.

```{r}
tolower(text[1])
toupper(text[1])
```

We can grab substrings with `substr`. The first argument is the string, the second is the beginning index (starting from 1), and the third is final index.

```{r}
substr(text[1], 1, 2)
substr(text[1], 1, 10)
```

This is useful when working with date strings as well:

```{r}
dates <- c("2015/01/01", "2014/12/01")
substr(dates, 1, 4) # years
substr(dates, 6, 7) # months
```

We can split up strings by a separator using `strsplit`. If we choose space as the separator, this is in most cases equivalent to splitting into words.

```{r}
strsplit(text[1], " ")
```

Let's dit into the data a little bit more. Given the construction of the dataset, we can expect that there will be many tweets mentioning relevant hashtags, such as #ICC, We can use the `grep` command to identify these. `grep` returns the index where the word occurs.

```{r}
grep('#ICC', text[1:10])
```

`grepl` returns `TRUE` or `FALSE`, indicating whether each element of the character vector contains that particular pattern.

```{r}
grepl('#ICC', text[1:10])
```

Going back to the full dataset, we can use the results of `grep` to get particular rows. First, check how many tweets mention the hashtag "#ICC".
```{r}
nrow(tweets)
grep('#ICC', tweets$text[1:10])
length(grep('#ICC', tweets$text))

```

It is important to note that matching is case-sensitive. You can use the `ignore.case` argument to match to a lowercase version.

```{r}
nrow(tweets)
length(grep('#ICC', tweets$text))
length(grep('#ICC', tweets$text, ignore.case = TRUE))
```

Another useful tool to work with text data is called "regular expression". You can learn more about regular expressions [here](http://www.zytrax.com/tech/web/regex.htm). Regular expressions let us develop complicated rules for both matching strings and extracting elements from them. 

For example, we could look at tweets that mention more than one hashtag using the operator "|" (equivalent to "OR")

```{r}
nrow(tweets)
length(grep('#ICC|#CPI', tweets$text, ignore.case=TRUE))
```

We can also use question marks to indicate optional characters.

```{r}
nrow(tweets)
length(grep('courts?', tweets$text, ignore.case=TRUE))
```

Other common expression patterns are:

- `.` matches any character, `^` and `$` match the beginning and end of a string.  
- Any character followed by `{3}`, `*`, `+` is matched exactly 3 times, 0 or more times, 1 or more times.  
- `[0-9]`, `[a-zA-Z]`, `[:alnum:]` match any digit, any letter, or any digit and letter.
- Special characters such as `.`, `\`, `(` or `)` must be preceded by a backslash.  
- See `?regex` for more details.

Another function that we will use is `gsub`, which replaces a pattern (or a regular expression) with another string:

```{r}
gsub('RT @[0-9_A-Za-z]+', 'USERNAME', text[1])
```

To extract a pattern, and not just replace, use parentheses and choose the option `repl="\\1"`:

```{r}
gsub('RT @([0-9_A-Za-z]+):.*', text[1], repl="\\1")

```

Now let's try to identify what tweets are probably not related to international courts. For example, how many tweets mention `cricket'?

```{r}
nrow(tweets)
length(grep('cricket', tweets$text, ignore.case=TRUE))
```

Let's assume we want to get rid of these. How would we do it? First, let's create a new column to the data frame that has value `TRUE` for tweets that mention this keyword and `FALSE` otherwise. Then, we can keep the rows with value `TRUE`.

```{r}
tweets$cricket <- grepl('cricket', tweets$text, ignore.case=TRUE)
table(tweets$cricket)
tweets.subset <- tweets[tweets$cricket==FALSE, ]
```
