---
title: "Scraping table data"
author: "Pablo Barbera"
date: "March 30, 2016"
output: html_document
---

### Scraping web data in table format

We will start by loading the `rvest` package, which will help us scrape data from the web.

```{r, message=FALSE}
#install.packages("rvest")
library(rvest)
```

The goal of this exercise is to scrape a table with the list of international courts from Wikipedia. 

The first step is to read the html code from the website we want to scrape, using the `read_html()` function. If we want to see the html in text format, we can then use `html_text()`.

```{r}
url <- 'https://en.wikipedia.org/wiki/International_court'
html <- read_html(url) # reading the html code into memory
html # not very informative
substr(html_text(html), 1, 1000) # first 1000 characters
```

To extract all the tables in the html code automatically, we use `html_table()`. Note that it returns a list of data frames, so in order to work with this dataset, we will have to subset the first element of this list.

```{r}
tab <- html_table(html)
str(tab)
icourts <- tab[[1]]
```

And we can now see what scope is more common

```{r}
sort(table(icourts$Scope))
```



